A scenario for a Planckian standard kilogram

### 2017-10-21 manuscript_gravitationalconstant.tex
2. A Description of a new G Measurement modified.

Historically, there have been several proposed unit systems that
included $G$ in some way,\cite{Tomilin1999} but the idea of including
both gravity and quantum in a unit system comes from
Planck\cite{Planck99}.  In the process of working towards the
radiation law, Planck realized that $c,G,h$ and the Boltzmann constant
$k_B$ (in his notation $G\rightarrow f$, $h\rightarrow b$,
$h/k_B\rightarrow a$) naturally define units for mass, length, time,
and temperature.  (Planck remarked that such units would be meaningful
even in non-human and extraterrestrial cultures.)

@@

The difficulty with Eq.~\eqref{eq:defs} is that all the expressions
involve $G$.  Given the current uncertainties in $G$,
Eq.~\eqref{eq:defs} would allow units to be standardized only to four
significant digits, which is obviously not enough.  Improvement by
several orders of magnitude would be needed, to make the Planckian
kilogram practicable.  In this paper, we study a scenario which may
make such an improvement possible.  A Planckian second not a prospect
with anything like present technology, because demands on the time
standard are more stringent still.

% In the late 19th century, Max Planck was working on reports about the irreversible process of radiation. He worked out a new entropy law to explain the empirical Wien law of radiation. This required him to introduce two new physical constants, which nowadays are known as the Planck constant and the Boltzmann constant . As a consequence, he described in 1899 a unit system which is based on fundamental natural constants only and thus free of anthropocentric arbitrariness. He pointed out the advantage of such a system in the following words:\cite{Planck99}\\

% \begin{quote} ...ihre Bedeutung f\"{u}r alle Zeiten und f\"{u}r alle, auch au{\ss}erirdische und au{\ss}ermenschliche Kulturen notwendig behalten und welche daher als  ``nat\"{u}rliche Ma{\ss}einheiten" bezeichnet werden k\"{o}nnen... \end{quote}

% \begin{quote} ...These necessarily retain their meaning for all times and for all civilizations, even extraterrestrial and non-human ones, and can therefore be designated as ``natural units"... \end{quote}

% Planck units are defined in such a way that five universal physical constants are given by a numerical value of 1 when expressed in these units. Today the five constants are the speed of light in vacuum $c$, the gravitational constant $G$, the reduced Planck constant $\hbar$, the Boltzmann constant $k_B$ and the Coulomb constant $\left(4 \pi \epsilon_0\right)^{-1}$. Planck had defined them in a somewhat different form and did not adopt any electromagnetic unit. But the natural extension is to include the Coulomb constant, another possibility is to use the elementary charge $e$ to define a basic unit of charge \cite{Tomilin99}.\\
% These constants can each be linked with a fundamental theory in physics that is widely accepted and universal: $\hbar$ with quantum mechanics, $c$ with special relativity, $G$ with general relativity, $k_B$ with thermodynamics and statistical mechanics, and $e$ or $\left(4 \pi \epsilon_0\right)^{-1}$ with electrodynamics.\\

% The International System of Units (SI) is the commonly used system of measurement. Its beginnings go back to the French Revolution in 1799 as a precursor of the metric system was implemented. With the signing of the Convention of the Metre in 1875, new international prototypes of the metre and the kilogram were established. In the early twentieth century electromagnetic units joined the system and it underwent constant improvement, until in 1960 the modern form, called the Système International d'Unités (SI), was launched. Since the Mole joined the system of measurement in 1971, the SI unit system is in its present form.\\

% To come away from the definitions based on artefacts, specific physical states and idealized experiments and because it is technologically possible to measure physical constants to high accuracy, the International Committee for Weights and Measures (CIPM) has proposed new definitions of the SI base units, which likely will be adopted in fall 2018 at the next General Conference on Weights and Measures. The revision of the SI unit system in 2018 goes towards a Planckian based system, since the values of seven physical constants get fixed and thereby define the seven base units. 
% To get a truly Planckian, hence ``most natural", basis for the SI, one would have to exchange the ground state hyperfine splitting frequency of the caesium-133 atom $\Delta \nu_{\mathrm{Cs}}$ with the gravitational constant $G$ and then define the second by  
% \begin{equation}
% \mathrm{s} = \sqrt{\frac{h \, G}{c^5}}.
% \end{equation}

% Because these four natural constants would have fixed numerical values, this definition is exact. Such a definition would lead to an SI system that is based on fundamental natural constants and therefore stable over time and realizable everywhere. Also completely new and more precise measurements could be designed without having to revise the definitions of the units \cite{Fischer16}. The big problem is, that the uncertainty in the measurements of $G$ is very large large. In Fig.~\ref{fig:Gmeasurements} the latest and most precise measurements of the gravitational constant are plotted together with the recommended CODATA value and its uncertainty. At first glance one notices, that the measured values are not even in good agreement with each other.\\

% The big problem that this approach faces, is that the current definition of the second which is based on the hyperfine splitting frequency in the Caesium 33 atom, is much more accurate. \\

@@

The measurements of the gravitational constant sketched in Fig.~\ref{fig:Gmeasurements} have all been done by measuring the very small force with a mechanical apparatus, expect the one with cold atom interferometry. However, it is difficult to exactly measure the weak gravitational force and it is doubtful if there will be significant improvements in measurement uncertainty with this techniques. Also the disagreement of current measurements suggests a demand for new, potentially different measurements. \\

\subsection{Description of new measurement}
Newton's law of gravitation does not only explain why an apple falls down from a tree, it also explains orbits. So if one has two bodies of masses $m_1$ and $m_2$, on which the only force acting is their respective gravitational attraction, they will orbit around their common barycenter. This orbit is described by Kepler's laws. The third law, given in Eq.~(\ref{eq:kepler3}), states that the square of the orbital period $P$ is directly proportional to the cube of the semimajor axis $a$ of the orbit. In this proportionality factor, the gravitational constant $G$ appears. To get a feeling what the dimensions of such an experiment could be, choose for example lead balls with a diameter of about 10 cm, their mass then is about 6 kg. If one then chooses a semimajor axis of 15 cm, the orbital period would be roughly 3.5 hours and the gravitational force acting between the two bodies is of order $10^{-8}$ N. In Fig.~\ref{fig:sketch} we show a sketch on how this would look like.
\begin{figure}
	\centering
	\includegraphics[width=0.9\textwidth]{img/sketchnonumb.pdf}
	\caption{Sketch on how the two masses inside the cavity of a spacecraft orbit around each other.}
	\label{fig:sketch}
\end{figure}
\begin{equation}\label{eq:kepler3}
P^2 = \frac{4 \, \pi^2}{G(m_1+m_2)}a^3
\end{equation}
If one measures the position of the two masses very accurately, it is possible to deduce the gravitational constant. The main difference to other techniques is that position measurements can be very accurate compared to directly measuring the gravitational force.\\

So the goal is to have a two-body system which is freely falling. The only way to achieve this over some time, is to go to space. Thus, one has to design a spacecraft with a cavity that contains the two masses and a position measurement apparatus. The spacecraft is then put in space, preferably to a Lagrange point of the earth-sun system, in order to prevent large inertial forces. To get an improvement in the precision of the gravitational constant, the positions of the two balls have to be known very accurately. Best suitable would probably be a laser interferometry system, which can yield precisions down to some nanometers.\cite{Loughridge13} \\

This approach appears simple, but it brings various difficulties with it. It is practically impossible to have a clean system, where only the respective gravitational force is acting between the two bodies, so one has find out what effects could possibly disturb the system and then either make them as small as possible or correct for them. \\

>>>>

Historically, there have been several proposed unit systems that
included $G$ in some way,\cite{Tomilin1999} but the idea of including
both gravity and quantum mechanics in a unit system comes from
Planck\cite{Planck99}.  In the process of working towards the
radiation law, Planck realized that $c,G,h$ and the Boltzmann constant
$k_B$ (in his notation $G\rightarrow f$, $h\rightarrow b$,
$h/k_B\rightarrow a$) naturally define units for mass, length, time,
and temperature.  (Planck remarked that such units would be meaningful
even in non-human and extraterrestrial cultures.)

@@

The difficulty with Eq.~\eqref{eq:defs} is that all the expressions
involve $G$.  Given the current uncertainties in $G$,
Eq.~\eqref{eq:defs} would allow units to be standardized only to four
significant digits, which is obviously not enough.  Improvement by
several orders of magnitude would be needed, to make the Planckian
kilogram practicable.  In this paper, we study a scenario which may
make such an improvement possible.  A Planckian second not a prospect
with anything like present technology, because demands on the time
standard are more stringent still.\\

@@

In Fig.~\ref{fig:Gmeasurements} we have plotted some of the most recent measurements of the gravitational constant. 3 of them have been carried out using a torsion balance (), Schlamminger et al. used a beam balance and ....... measured the separation of two simple pendulums in a gravitational field. They all tried to measure the very weak gravitational force with a mechanical apparatus. The sixth measurement was performed under another approach, using cold atom interferometry. Its uncertainty is a lot larger than the others, but it could become more competitive in the future. However, a brief glance on the plotted measurements shows the disagreement of current $G$-values and suggests a demand for new, potentially different techniques.\\

\subsection{Description of new measurement}
Newton's law of gravitation does not only explain why an apple falls down from a tree, it also explains orbits. If one has two bodies of masses $m_1$ and $m_2$, on which the only force acting is their respective gravitational attraction, they will orbit around their common barycenter in an ellipse described by Kepler's laws. The third law, given in Eq.~(\ref{eq:kepler3}), states that the square of the orbital period $P$ is directly proportional to the cube of the semimajor axis $a$. In this proportionality factor, the gravitational constant $G$ appears.
\begin{equation}\label{eq:kepler3}
P^2 = \frac{4 \, \pi^2}{G(m_1+m_2)}a^3
\end{equation}
If one can measure the position of the two masses very accurately and thus knows the orbit, it is possible to deduce the gravitational constant. 
So what we need is a freely falling two-body system. The only way to achieve this over some time, is to go to space. Thus, one has to design a spacecraft involving a cavity that contains the two masses and a position measurement apparatus. The cavity should shield the two-body system from environmental influences. The spacecraft is then put in space, preferably to a Lagrange point of the earth-sun system, in order to prevent large inertial forces.\\

\begin{figure}
	\centering
	\includegraphics[width=0.9\textwidth]{img/sketchnonumb.pdf}
	\caption{Sketch on how the two masses inside the cavity of a spacecraft orbit around each other.}
	\label{fig:sketch}
\end{figure}

To get a feeling what the dimensions of such an experiment could be, choose for example lead balls with a diameter of about 10 cm, their mass then is about 6 kg. If one then chooses a semimajor axis of 15 cm, the orbital period would be roughly 3.5 hours and the gravitational force acting between the two bodies is of order $10^{-8}$ N. In Fig.~\ref{fig:sketch} we show a sketch on how this would look like.
For getting an improvement in the precision of the gravitational constant, the positions of the two balls have to be known very accurately. Best suitable would probably be a laser interferometry system, which can yield precisions down to some nanometers.\cite{Loughridge13}
The main difference to other techniques is that position measurements can be very accurate compared to directly measuring the gravitational force.
This new kind of measurement of the gravitational constant appears simple, but it brings various difficulties with it. It is practically impossible to have a clean system, where only the respective gravitational force is acting between the two bodies, so one has to find out what effects could possibly disturb the system and then either make them as small as possible or correct for them. \\
